Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2017-10-15 TACAS2018/paper.tex
Changes with cristina

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Verification of Continuous time systems}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \todo{I find this section header unclear. What is being verified here? Stability and safety only come later on.}
 The main problem with the above discretization is that is does not allow us to observe the system at arbitrary points in time, only at integer multiples of the sample time. While this is adequate for the controller, it is not sufficient for verification. We must therefore establish a new model that allows us to observe the system at a given time $t$.
 %We achieve this by using acceleration as described below.

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsubsection{Discrete dynamics using reals for systems without inputs}\label{sec:real_discrete_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \begin{theorem}
 Given $\dot{\vec{x}}=\mat{A}\vec{x}$ \todo{$\dot{\vec{x}}(t)$?}, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, the expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  $\vec{x}_T$ is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x} : k \in (0,\infty)$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
% \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsubsection{Discrete dynamics using reals for systems with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}+(\mat{I}-\mat{A}_d^k\mat{B}\vec{u}) : k \in (0,\infty)$ is a witness of the system at time $kT_d$.
 \end{corollary}
% \begin{proof}
% We begin by expanding the equation
% \begin{align}
% \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\nonumber\\
% &=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u}\nonumber\\
% &= \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}
% \end{align}
% where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
% \end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsubsection{Discrete dynamics using reals for systems with discrete time feedback inputs}\label{sec:real_discrete_feedback_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Time delay on discrete time feedback systems} \label{sec:delay}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\todo{How does fit in the description at the beginning of sec 3.3?}

A common ocurrence in digital systems is that they do not necessarily
satisfy the real-time specifications.  Delays due to processing time,
communication lines, or other mechanisms will cause disturbances in the
control loop.  In our model these delays are all deferred to the ZOH, i.e.,
we presume that everything that happens in the digital world takes place at
the sampling instance $t$ and the output of the result is delayed at the DAC
when it is fed back into the real world.  We~will also presume that this
delay is much smaller than the sampling time.

\todo{equations \eqref{eq:feedback_sampled_cont} and
  \eqref{eq:feedback_cont} are in the appendix so it's preferable to not
refer to them so directly.}
Let us look at equations \eqref{eq:feedback_sampled_cont} and
\eqref{eq:feedback_cont}.  When we introduce an arbitrary delay $T_d$ where
$t_d=T_d$ and $t_s=T_s-T_d$ we obtain:
%
\begin{align}
\vec{x}_{k} &=\mat{A}_{t_s}(\mat{A}_{t_d}\vec{x}_{k-1}-\mat{B}_{t_d}\mat{K}\vec{x}_{k-2})-\mat{B}\mat{K}\vec{x}_{k-1}\nonumber\\
&=  (\mat{A}_{T_s}-\mat{B}_{t_s}\mat{K})\vec{x}_{k-1}-\mat{A}_{t_s}\mat{B}_{t_d}\mat{K}\vec{x}_{k-2}
\label{eq:delay_cont}
\end{align}
%
Differentiating with the expected progression without delay, we get a delay
error:
%
\begin{align}
{\vec{x}_e}_{k}=&(\mat{B}_{T_s}-\mat{B}_{t_s})\mat{K}\vec{x}_{k-1}-\mat{A}_{t_s}\mat{B}_{t_d}\mat{K}\vec{x}_{k-2}\nonumber\\
=&(\mat{B}_{T_s}-\mat{B}_{t_s})\vec{u}_{k}-\mat{A}_{t_s}\mat{B}_{t_d}\vec{u}_{k-1}
\end{align}
Rather than explicitly calculate this error, we will find an upper bound. Let $|U|_\infty=max(|\vec{u}|_\infty \mid \vec{u} \in U)$ be the maximum infinity norm of a vector set, propagating the error to the output $\vec{y}_k$ we have:
\begin{align}
|{\vec{y}_e}_{k}|_1 \leq q_t=&(|\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})|_1+|\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}|_1)|U|_\infty
\end{align}

We overapproximate the effects of this vector by introducing a noise signal at the output $\nu_t=[-q_t\ q_t]$ and use the dynamics of the system without delay.

@@

The models we consider have two sources of error that are due to numerical  
representation.  The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$.
The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{sec:numeric_rep}.

@@

where $0\leq p_s \leq 1$ is the desired percentage of the target within \todo{typo?} to remain after $t_s$.

@@

As stated in Section~\ref{ssec:stabspecification}, we use Jury's Criterion
to assert this specification. \todo{Not clear to me what ``this specification'' refers to.}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Stability of a Digitally Observed System}\label{sec:cof_fwl_stability}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

%% Now that we have established convergence, we may examine the safety
%% specification of the state space.
In the case of SISO systems, the safety specification is
given by a range of the input and a range of the output which are considered
safe ($\vec{u}_k \in [\underline{u}\ \ \overline{u}] \wedge \vec{y}_k \in
[\underline{y}\ \ \overline{y}]$)
\todo{I think the safety specification is generally on the states of the system.}.
Whilst there are other properties we may
include in the verification, we will solely focus on these for simplicity of
the explanation.  We assume that we do not want saturation in our
system, hence the control loop has only one mode.
%We have also identified the quantization noise as a bounded non-deterministic input.
First we must
translate the safety specification into the statespace.  Replacing on the
second part of equation \eqref{eq:observer_LTI_cf} we obtain
%
\begin{equation}
\phi_{safety} := \underline{y}<\mat{C}\vec{x}_k<\overline{y}, \forall k \geq 0
\label{eq:of_bounds}
\end{equation}
\todo{what happened to u when going from \eqref{eq:observer_LTI_cf}?}
%
%Note that we use the bounds of the input after the transformation by $\mat{B}$ since the matrix may reverse the bounds. 
%
\todo{I thought \eqref{eq:of_bounds} was the safety spec. Why is it solved?}
Equation~\eqref{eq:of_bounds} can be solved using linear programming as it
describes a polyhedral region in $\mathbb{R}^p$.  The second \todo{second?} step is
evaluating the progression of the system over time in order to calculate
$\vec{x}_k$ to validate against this specification.  For this step we require
an initial set $\vec{x}_0$ from which the system progresses.  Obviously, this set
%$\vec{x}_0$
must satisfy the specification, otherwise the system will be
unsafe to begin with.  We will accept a specification in the form
$\mat{E}\mat{T}\vec{x}_0<\mat{f}$ \todo{ET? f?}.  The presence of $\mat{T}$ is because
this will typically be given in the original state-space.

@@

\item 
The {\sc ABSTRACT} module starts by creating an abstraction from the initial specification,
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bitvector size is required for
the synthesis process.
\item Build the Observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in section \ref{sec:numeric_rep}.
\item Transform the system into Canonical form and add nondeterminism for the errors
using the techniques described in section \ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in section \ref{ssec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in sections \ref{sec:delay} and \ref{sec:quantization}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.  
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} x_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
The Abstraction is fed to the synthesiser.
\item In the {\sc synthesize} phase, we synthesize a candidate controller $\mat{K} \in \mathbb{R}_{\langle C\rangle}^p$
and observer $\mat{L} \in \mathbb{R}_{\langle C\rangle}^p$ that satisfy
$\phi_\mathit{stability} \wedge \phi_\mathit{safety} \wedge \phi_\mathit{init}^{K}$ by invoking a SAT solver.
There are two exit conditions:
\item If there is no candidate solution we return UNSAT and signal the modeller. Based on the current abstraction,
the {\sc  model} phase may choose to increase the precision, decrease the sample time, or return a synthesis failure.
\item If we have a candidate solution we will feed it to the verifier in order to validate it.
\item The verifier works in multiple stages.
\item The {\sc Verify Synthesis} submodule verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy the tighter
specification, it feeds back to the modeller which uses sensitivity analysis to
synthesise a refined solution. 
Formally speaking, given a violating controller $\{ \mat{K},\mat{L} \} \not \models \phi_{static}$
we look for a variation $\mat{\delta}_K, \mat{\delta}_L$ such that
$\{ \mat{K}', \mat{L}' \}=\{ \mat{K}+\mat{\delta}_K,\mat{L}+\mat{\delta}_L \} \models \phi_{static}$, and similarly for the observer.
%\mat{K}' \in \mathbb{R}_{\langle C \rangle}$. 
If the Modeller cannot synthesise a suitable candidate($\mat{K}' \not
\models \phi_{static}^K$), it returns to the synthesiser with a request for
either higher precision or a lower upper bound $\overline{\lambda}$. Otherwise
the refined candidate solution is given to the second verification stage.
\item The {\sc Discrete Time} submodule verifies $\phi_{safety}$ using discrete
time dynamics.  If the verification fails, the verifier reports to the

>>>>

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of Continuous time systems}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \todo{I find this section header unclear. What is being verified here? Stability and safety only come later on.}
 The main problem with the above discretization is that is does not allow us to observe the system at arbitrary points in time, only at integer multiples of the sample time. While this is adequate for the controller, it is not sufficient for verification. We must therefore establish a new model that allows us to observe the system at a given time $t$.
 %We achieve this by using acceleration as described below.

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelrated dynamics using reals for systems without inputs}\label{sec:real_discrete_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 The state $\vec{x}_T$ present at time $T$ can be expressed as a linear function of $\vec{x}_0$ as described below
 \begin{theorem}
 Given $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ \todo{$\dot{\vec{x}}(t)$?}, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, the expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  $\vec{x}_T$ is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0 : k \in (0,\infty)$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
% \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics using reals for systems with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0+(\mat{I}-\mat{A}_d^k\mat{B}\vec{u}) : k \in (0,\infty)$ is a witness of the system at time $kT_d$.
 \end{corollary}
% \begin{proof}
% We begin by expanding the equation
% \begin{align}
% \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\nonumber\\
% &=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u}\nonumber\\
% &= \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}
% \end{align}
% where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
% \end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics using reals for systems with discrete time feedback inputs}\label{sec:real_discrete_feedback_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The final case we analyse here is that of hybrid time systems, where the
plant dynamics evolve in continuous time while the feedback dynamics evolve
in discrete time, i.e. a new input value is only produced for every sample interval.

\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T, \mat{B}_T$ as per \eqref{eq:continuous_tube_dyn} and $\mat{A}_d, \mat{B}_d$ as per equation \eqref{eq:discretize} is a witness of the system
 \begin{align}
 \dot{\vec{x}}(t) &= \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\nonumber\\
 \vec{u}_t&=\mat{K}\vec{x}_k  \mid  kT_s \leq t < (k+1)T_s
 \end{align}
 \end{theorem}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Time delay on discrete time feedback systems} \label{sec:delay}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
A common ocurrence in digital systems is that they do not necessarily
satisfy the real-time specifications.  Delays due to processing time,
communication lines, or other mechanisms will cause disturbances in the
control loop.  In our model these delays are all deferred to the ZOH, i.e.,
we presume that everything that happens in the digital world takes place at
the sampling instance $t$ and the output of the result is delayed at the DAC
when it is fed back into the real world.  We~will also presume that this
delay is much smaller than the sampling time.

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upperbound by 
\begin{align}
|{\vec{y}_e}_{k}|_1 \leq q_t, \text{ where } q_t=&(|\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})|_1+|\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}|_1)|U|_\infty
\end{align}
\end{theorem}

@@

The models we consider have two sources of error that are due to numerical  
representation which we commonly refer to as the FWL effect.
The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$.
The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{sec:numeric_rep}.

@@

where $0\leq p_s \leq 1$ is the desired percentage of the target within which the output must remain after $t_s$.

@@

As stated in Section~\ref{ssec:stabspecification}, we use Jury's Criterion on these polynomials
to assert the stability specification.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Stability of an Observed System with FWL effects}\label{sec:cof_fwl_stability}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

%% Now that we have established convergence, we may examine the safety
%% specification of the state space.
In the case of SISO systems, the safety specification is
given by a range of the input and a range of the output which are considered
safe ($\vec{u}_k \in [\underline{u}\ \ \overline{u}] \wedge \vec{y}_k \in
[\underline{y}\ \ \overline{y}]$)
\todo{I think the safety specification is generally on the states of the system.}.
Whilst there are other properties we may
include in the verification, we will solely focus on these for simplicity of
the explanation.  We assume that we do not want saturation in our
system, hence the control loop has only one mode.
%We have also identified the quantization noise as a bounded non-deterministic input.
First we must
translate the safety specification into the statespace.  Replacing on the
second part of equation \eqref{eq:observer_LTI_cf} we obtain
%
\begin{equation}
\phi_{safety} := \underline{y}<\mat{C}\vec{x}_k<\overline{y}, \forall k \geq 0
\label{eq:of_bounds}
\end{equation}
%
%Note that we use the bounds of the input after the transformation by $\mat{B}$ since the matrix may reverse the bounds. 
%
%We evaluate the progression of the system over time by calculating
%$\vec{x}_k$ to validate against this specification.  For this step we require
%an initial set $\vec{x}_0$ from which the system progresses.  Obviously, this set
%must satisfy the specification, otherwise the system will be
%unsafe to begin with.  For efficiency purposes, the inequality in \eqref{eq:of_bounds} is transformed into reachable form using $\mat{T}_c, \mat{T}_o$ which reduces the complexity from $\mathcal{O}(p^2)$ to $\mathcal{O}(p)$.

@@

\begin{enumerate}
\item 
The {\sc ABSTRACT} module starts by creating an abstraction from the initial specification,
\begin{enumerate}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bitvector size is required for
the synthesis process.
\item Build the Observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in section \ref{sec:numeric_rep}.
\item Transform the system into Canonical form and add nondeterminism for the errors
using the techniques described in section \ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in section \ref{ssec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in sections \ref{sec:delay} and \ref{sec:quantization}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
%
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} x_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bitvectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{enumerate}
There are two abstractions created by these specifications: $\phi_{synth}$ which is given to the synthesiser and $\phi_{static}$ which is given to the synthesis refinement loop. These abstractions will be refined independently, so we will not detail what each will contain at a given stage. We only note that at all times $\phi_{static} \models \phi_{synth}$.
